Combined stations and predators

### 2023-10-26 reports/03_combined.qmd
Add supplemental table 1

tar_load(c("predators_agg", "predators_stations", "zoop", "ant_sf"), 
         store = here::here("_targets"))
source(here::here("R", "sp.R"))
source(here::here("R", "zoop.R"))
```

@@

Each station was associated with the modal ice type (multi-year, first-year, thin, or open) from the underway observations, which were only associated with predator sightings. **Seems weird that no stations shook out as multi-year ice, right?** This figure shows how frequent the modal ice type was at each station. The two grey lines represent 1:1 and 2:1. Stations falling on the upper line had *uniform* ice category observations i.e., every interval within 15 km had the same ice category. Stations falling below the lower line had *highly mixed* ice category observations i.e., the most frequent ice category was observed at less than half the intervals within 15 km of the station. Most points on the figure include multiple stations, indicated by the color. For example, in the THIN ice category (bottom), 8 stations (red point) had 4 intervals with ice category observations where all 4 were THIN (x,y = 4,4).

@@

predators_stations %>% 
  mutate(n_obs = ice_intervals,
         n_mode = round(ice_type_freq * n_obs, digits = 0)) %>% 
  group_by(n_obs, n_mode, ice_type) %>% 
  summarize(Stations = n(), .groups = "drop") %>% 
  ggplot(aes(n_obs, n_mode, color = Stations)) +
  geom_abline(slope = c(1, 0.5), intercept = 0, color = "grey50") +
  geom_point() +
  labs(x = "Count of intervals with ice category observation",
       y = "Count of intervals with most frequent ice category") +
  scale_color_fermenter(palette = "RdYlBu",
                        breaks = c(10, 20, 100, 200)) +
  coord_fixed() +
  facet_wrap(~ ice_type, ncol = 2) +
  theme_classic()
```

@@

ant_basemap(map_lim) +
  geom_sf(aes(color = ice_type), 
          filter(predators_stations, year != 2014),
          size = 0.5) +
  scale_color_brewer("Ice category", palette = "YlGnBu") +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  facet_wrap(~ year) +
  theme(legend.position = "left")

@@

>>>>

tar_load_everything(store = here::here("_targets"))
source(here::here("R", "sp.R"))
source(here::here("R", "zoop.R"))
```

@@

```{r}
ice_qa <- zoop_sf %>% 
  right_join(stations_ice, by = "amlr.station") %>%
  distinct(amlr.station, ice_type, ice_intervals, ice_type_freq) %>%
  mutate(n_obs = ice_intervals,
         n_mode = round(ice_type_freq, digits = 0))
```

Each station was associated with the modal ice type (multi-year, first-year, thin, or open) from the underway observations. This figure shows how frequent the modal ice type was at each station. The two grey lines represent 1:1 and 2:1. Stations falling on the upper line had *uniform* ice category observations i.e., every interval within 15 km had the same ice category. Stations falling below the lower line had *highly mixed* ice category observations i.e., the most frequent ice category was observed at less than half the intervals within 15 km of the station.

The frequency of the modal ice type was 90% or greater for 57.8% of stations. The modal ice type was \<50% frequency at just 10 stations (\<5% of total stations).

@@

ggplot(ice_qa, aes(n_obs, n_mode)) +
  geom_abline(slope = c(1, 0.5), intercept = 0, color = "grey50") +
  geom_point(alpha = 0.5) +
  labs(x = "Count of intervals with ice category observation",
       y = "Count of intervals with most frequent ice category") +
  coord_fixed() +
  facet_wrap(~ ice_type, ncol = 2) +
  theme_classic()
```

@@

ice_sf <- station_effort %>% 
  select(amlr.station, year) %>% 
  right_join(stations_ice, by = "amlr.station")

ant_basemap(map_lim) +
  geom_sf(aes(color = ice_type), 
          ice_sf,
          size = 0.5) +
  scale_color_brewer("Ice category", palette = "YlGnBu") +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  facet_wrap(~ year) +
  theme(legend.position = "left")

@@

```{r}
ice_type_grand_total_row <- stations_ice %>% 
  mutate(YEAR = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
  count(ice_type) %>% 
  pivot_wider(names_from = "ice_type", 
              values_from = "n", 
              values_fill = 0) %>% 
  mutate(TOTAL = OPEN + THIN + FY + MY,
         across(-TOTAL, \(x) x / TOTAL),
         YEAR = "Total")

stations_ice %>% 
  mutate(YEAR = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
  count(YEAR, ice_type) %>% 
  pivot_wider(names_from = "ice_type", 
              values_from = "n", 
              values_fill = 0) %>% 
  mutate(TOTAL = OPEN + THIN + FY + MY,
         across(c(OPEN, THIN, FY, MY), \(x) x / TOTAL)) %>% 
  rbind(ice_type_grand_total_row) %>% 
  mutate(across(-c(TOTAL, YEAR), scales::percent)) %>% 
  knitr::kable()

stations_ice %>% 
  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
  count(year, ice_type) %>% 
  group_by(year) %>% 
  mutate(frac = n / sum(n)) %>% 
  ungroup() %>% 
  mutate(ice_type = fct_rev(ice_type)) %>% 
  ggplot(aes(year, frac, fill = ice_type)) +
  geom_col() +
  scale_fill_brewer(palette = "Blues") +
  scale_y_continuous("Sea ice composition", labels = scales::percent) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank())
stations_ice %>% 
  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
  group_by(year, ice_type) %>% 
  summarize(mean_cov = mean(ice_coverage) / 10,
            n = n(),
            .groups = "drop_last") %>% 
  mutate(n = sum(n),
         se_cov = sqrt(mean_cov * (1 - mean_cov) / n)) %>% 
  ungroup() %>% 
  mutate(ice_type = fct_rev(ice_type)) %>% 
  ggplot(aes(year, mean_cov, fill = ice_type)) +
  geom_errorbar(aes(ymin = mean_cov - se_cov, ymax = mean_cov + se_cov),
                width = 0.2,
                color = "grey50",
                position = "dodge") +
  geom_point(size = 2, 
             shape = 21, 
             color = "grey50", 
             position = position_dodge(width = 0.2)) +
  scale_fill_brewer(palette = "Blues") +
  scale_y_continuous(limits = c(NA, 1),
                     labels = scales::percent) +
  theme_minimal() +
  theme()

stations_ice %>% 
  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
  group_by(year) %>% 
  summarize(mean = mean(ice_coverage),
            sd = sd(ice_coverage),
            .groups = "drop") %>% 
  mutate(mean_sd = str_glue("{format(mean, digits = 2)} \U00B1 {format(sd, digits = 2)}")) %>% 
  knitr::kable()
```

## Effort

### Number of stations

Total of `r n_distinct(predators_stations$amlr.station)` stations. Breakdown by year:

```{r}
predators_stations %>% 
  distinct(amlr.station) %>% 
  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
  count(year)
```

### Line transects

```{r}
survey_nmi <- predators_stations %>% 
  distinct(amlr.station, survey_nmi) %>% 
  pull(survey_nmi)
mean_nmi <- format(mean(survey_nmi), digits = 3) 
sd_nmi <- format(sd(survey_nmi), digits = 3) 
```

`r mean_nmi` $\pm$ `r sd_nmi` nm of survey effort associated with each station.

### Ice and effort combined

```{r}
percent <- partial(scales::percent, accuracy = 0.1)

ice_type_lookup = c(
  OPEN = "Open water",
  THIN = "Thin ice",
  FY = "First-year ice",
  MY = "Multi-year ice"
)

ice_table <- stations_ice %>% 
  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
  group_by(year, ice_type) %>% 
  summarize(n_year_type = n(),
            ice_mean = mean(ice_coverage / 10),
            .groups = "drop_last") %>% 
  mutate(n_year = sum(n_year_type),
         ice_se = sqrt(ice_mean * (1 - ice_mean) / n_year)) %>% 
  ungroup() %>% 
  transmute(year,
            ice_type = ice_type_lookup[ice_type],
            ice_val = str_glue("{n_year_type} ({percent(ice_mean)} \U00B1 {percent(ice_se)})")) %>% 
  pivot_wider(names_from = ice_type, 
              values_from = ice_val, 
              values_fill = "0") %>% 
  rbind(
    stations_ice %>% 
      mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
      group_by(ice_type) %>% 
      summarize(n_type = n(),
                ice_mean = mean(ice_coverage / 10)) %>% 
      mutate(ice_se = sqrt(ice_mean * (1 - ice_mean) / sum(n_type)),
             ice_val = str_glue("{n_type} ({percent(ice_mean)} \U00B1 {percent(ice_se)})")) %>% 
      transmute(year = "Total", ice_type = ice_type_lookup[ice_type], ice_val = str_glue("{n_type} ({percent(ice_mean)} \U00B1 {percent(ice_se)})")) %>% 
      pivot_wider(names_from = ice_type, 
              values_from = ice_val)
  )

effort_table <- predators_stations %>% 
  distinct(amlr.station, survey_nmi) %>% 
  mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
  group_by(year) %>% 
  summarize(`Stations sampled` = n(),
            mean_nmi = mean(survey_nmi),
            sd_nmi = sd(survey_nmi),
            `Effort per station (nmi)` = str_glue("{sprintf('%0.1f', mean_nmi)} \U00B1 {sprintf('%0.1f', sd_nmi)}")) %>% 
  select(year, `Stations sampled`, `Effort per station (nmi)`) %>% 
  rbind(
    predators_stations %>% 
      distinct(amlr.station, survey_nmi) %>% 
      mutate(year = str_extract(amlr.station, "AMLR([0-9]{4}).*", 1)) %>% 
      summarize(
        year = "Total",
        `Stations sampled` = n(),
        `Effort per station (nmi)` = str_glue("{sprintf('%0.1f', mean(survey_nmi))} \U00B1 {sprintf('%0.1f', sd(survey_nmi))}")
      )
  )

effort_table %>% 
  left_join(ice_table, by = "year") %>% 
  knitr::kable() %>% 
  kableExtra::add_header_above(c(" ", "Survey effort" = 2, "Ice conditions" = 4))
```
